Multiscale Modeling of the Paradoxical Roles of the Immune System in EMT-Mediated Cancer

### 2019-01-12 Paper/Paper.tex
Included Adam's Edits and Began My Responses
an ad hoc merging of Adam’s edits and the beginnings of my responses

[No subsections headings within Introduction]
\subsection{The Tumor}\label{TheTumor}
The tumor influences the immune system by its presence and by the cellular processes it engages in that result in shaping the TME.
On the one hand, the mutated DNA of tumor cells results in these cells being recognized as foreign by members of the innate immune system such as natural killer (NK) cells and dendritic cells.
It is then these cells that initiate the cascade of cytokines that result in the full activation of the body's immune system to fight the non-self cells.
On the other hand, the tumor cells actively work to shape the immune response that the body launches.
This is accomplished through its own set of cytokines, such as transforming growth factor beta (TGF-$\beta$), that work to create a tumor-supportive environment, including shifting the immune balance to be immunosuppressive by enhancing the recruitment of Tregs. Since TGF-$\beta$ is also a driver of EMT, then through direct release of TGF-$\beta$ and recruitment of TGF-$\beta$-secreting Treg cells, cells of the tumor and the surrounding tissue are driven towards a more mesenchymal profile.

@@

The research on mesenchymal cells' immune evasiveness is quite conclusive with the research even going so far as establishing functional causes for this phenomenon.
Once cytotoxic immune cells, for example natural killer cells and cytotoxic T cells, have targeted a cell to lyse, they necessarily must make a physical connection with that cell, termed an immunological synapse. Ref 
However, due to down-regulated surface markers on mesenchymal cells, they are refractory to this immunological task thus rendering them less susceptible to immune clearance. Ref 
In line with known motility of mesenchymal cells, scientists have referred to this condition as mesenchymal cells being ``slippery.''
This quality of mesenchymal cells is then quantified in our model by the parameter, MIE, for mesenchymal immune evasion.

@@

Also, all proliferating cells have a probability of undergoing a pathway mutation.
If they do, one pathway is randomly chosen among the three and is mutated.
If the cell does not undergo a mutation, then its probability of mutating during subsequent cell increases. [**]

@@

\subsection{Sensitivity Analysis}\label{SensAnalysis}
To assess the sensitivity to the many model parameters [how many?], a Morris OAT algorithm was implemented. [need to define] Ref 
The parameters were all given Gaussian distributions centered at their initial value and a large standard deviation.
Those that needed bounds were truncated.
The results of the Morris OAT can be found in Figure \ref{fig:ModelIntro}B.
Notice that the following parameters have the most global influence over the system. [which parameters?]
While it makes sense for such and such [??] parameters to have large influences over patient outcomes, these fall outside of the scope of the current paper and instead it is preferable to focus on the following subset.

\subsection{Mesenchymal Cell Properties Dramatically Change Survival Outcomes}\label{MesPars}
The two quantities that change when a cell transitions to a mesenchymal state are its immune evasion and its increased quiescence, MIE and MQ, respectively.
Also involved in the EMT process, is the cytokine TGF-$\beta$, which upregulates EMT.
By the Morris OAT analysis presented in Section \ref{SensAnalysis}, these three parameters do have large impacts on survival outcomes and each warrant exploration.

@@

\begin{figure}[H]
\center
\frame{\includegraphics[width=\columnwidth]{Figure3/Figure3.pdf}}
\caption{A.Survival curves for various inflammation cycling schemes. In the top two plots, the Time to Cancer decreases as MIE increases. In the bottom plot, there is no significant difference in survival.
B. The average Time to Cancer for each cohort.
C. Survival curves for various inflammation cycling schemes. In the top two plots, the Time to Cancer reaches a maximum value somewhere near MQ = 0.3. In the bottom plot, there is no significant difference in survival.
D. The average Time to Cancer for each cohort.}
\label{fig:VaryINFL_and_MesPars}
\end{figure}

>>>>

[No subsections headings within Introduction]
\subsection{The Tumor}\label{TheTumor}
The tumor influences the immune system by its presence and by the cellular processes it engages in that result in shaping the TME.
On the one hand, the mutated DNA of tumor cells results in these cells being recognized as foreign by members of the innate immune system such as natural killer (NK) cells and dendritic cells.
It is then these cells that initiate the cascade of cytokines that result in the full activation of the body's immune system to fight the non-self cells.
On the other hand, the tumor cells actively work to shape the immune response that the body launches.
This is accomplished through its own set of cytokines, such as transforming growth factor beta (TGF-$\beta$), that work to create a tumor-supportive environment, including shifting the immune balance to be immunosuppressive by enhancing the recruitment of Treg cells. Since TGF-$\beta$ is also a driver of EMT, then through direct release of TGF-$\beta$ and recruitment of TGF-$\beta$-secreting Treg cells, cells of the tumor and the surrounding tissue are driven towards a more mesenchymal profile.

@@

The research on mesenchymal cells' immune evasiveness is quite conclusive with the research even going so far as establishing functional causes for this phenomenon.
Once cytotoxic immune cells, for example NKs and CTLs, have targeted a cell to lyse, they necessarily must make a physical connection with that cell, termed an immunological synapse. Ref 
However, due to down-regulated surface markers on mesenchymal cells, they are refractory to this immunological task thus rendering them less susceptible to immune clearance. Ref
In line with known motility of mesenchymal cells, scientists have referred to this condition as mesenchymal cells being ``slippery.''
This quality of mesenchymal cells is then quantified in our model by the parameter, MIE, for mesenchymal immune evasion.

@@

Also, all proliferating cells have a probability of undergoing a pathway mutation.
If they do, one is randomly chosen among the three and that one either mutates or remains mutated.
If they do, one pathway is randomly chosen among the three and is mutated.
If the cell does not undergo a mutation, then its probability of mutating during subsequent cell increases. [**]

@@

\subsection{Sensitivity Analysis}\label{SensAnalysis}
To assess the sensitivity to the many model parameters [how many?], a Morris OAT algorithm was implemented. [need to define] Ref 
The parameters were all given Gaussian distributions centered at their initial value and a large standard deviation.
Those that needed bounds were truncated.
The results of the Morris OAT can be found in Figure \ref{fig:ModelIntro}D.
Clearly, a certain subset of parameters have an outsized effect on the model while others do relatively little.
The two most influential according to this analysis are related to Treg cells: their death rate and recruitment.
This makes sense as Treg cells both suppress cytotoxic effects of other immune cells and also secrete significant quantities of TGF-$\beta$ which drives EMT.
Thus, Treg cells are intimately tied to all three aspects of the model and so their presence or lack thereof significantly effects the system.
However, we want to better isolate the effects of EMT on these dynamics, so parameters such as MIE and MGA are of interest.
In addition, inflammation parameters dictating the cycling scheme are interesting.
What we can see from this plot is that the model also has some high sensitivity to these parameters.
In terms of Treg cells, their secretion of TGF-$\beta$ is highly significant and so we will also look at that parameter.

\subsection{Mesenchymal Cell Properties Dramatically Change Survival Outcomes}\label{MesPars}
The two quantities that change when a cell transitions to a mesenchymal state are its immune evasion and its increased quiescence, MIE and MGA, respectively.
Also involved in the EMT process, is the cytokine TGF-$\beta$, which upregulates EMT.
By the Morris OAT analysis presented in Section \ref{SensAnalysis}, these three parameters do have large impacts on survival outcomes and each warrant exploration.

@@

\begin{figure}[H]
\center
\frame{\includegraphics[width=\columnwidth]{Figure3/Figure3.pdf}}
\caption{A.Survival curves for various inflammation cycling schemes. In the top two plots, the Time to Cancer decreases as MIE increases. In the bottom plot, there is no significant difference in survival.
B. The average Time to Cancer for each cohort.
C. Survival curves for various inflammation cycling schemes. In the top two plots, the Time to Cancer reaches a maximum value somewhere near MGA = 0.3. In the bottom plot, there is no significant difference in survival.
D. The average Time to Cancer for each cohort.}
\label{fig:VaryINFL_and_MesPars}
\end{figure}
